Research Notes

### 2018-12-30 writeups/notes.tex

\subsection{Gotcha: $\nu_x, \nu_z$ and Grid Spacing}

If the grid resolution is uneven in the $x, z$ directions, it might be tempting
to use different $\nu_x, \nu_z$ per the above criteria, since their $k_{turb} =
k_{\max}$ differ. This is dangerous however, since $\nu \nabla^2 \vec{u}_1$ has
to be divergenceless! Thus, it is more advisable to choose an equal grid spacing
in the two directions.

>>>>

\subsection{Choices of $k_x, k_z$ and Scales}

Lots of scales to consider:
\begin{itemize}
    \item $k_zH \lesssim 1$ is required for the density stratification to be
        considered weak.

    \item $k_z \ll k_{z, \max} \equiv \frac{2\pi N_z}{L_z}$ is required for
        there to be significant separation of scales between the linear mode,
        the nonlinear cascade and the grid resolution.

    \item $L_Z \gtrsim 10H$ or a few, to get a decent amplitude growth going
        from linear to nonlinear amplitudes.

    \item Viscosity induced decay length scale: $\partial_t \sim \omega + i\nu
        k^2$ where $k^2 = k_x^2 + k_z^2$, so the characteristic length scale on
        which the wave decays is $L_\nu \sim \frac{v_{ph, z}}{\nu k_z^2}$. Call
        $\tilde{\nu} = \nu / \frac{\omega}{k_zk_{z, \max}}$, \lstinline{NU_MULT}
        in our simulations, then we may simplify
        \begin{equation}
            L_\nu \sim \frac{k_z^2k_{z, \max}}{\tilde{\nu}k^4}.
        \end{equation}
        Approximating $k_z \approx k$ and requiring $L_\nu \gtrsim L_Z$ enforces
        \begin{equation}
            \frac{L_\nu}{L_Z} \sim \frac{N_z}{\tilde{\nu}200\pi} \gtrsim 1.
        \end{equation}
\end{itemize}

However, in order for the simulation to both be well-damped
at grid resolutions and be neglibily damped at leading order nonlinear
wavelengths, we require $k_{0z} \ll \frac{1}{L_{NL}} \ll \frac{2\pi
N_z}{L_z}$, where $L_{NL}$ denotes some characteristic length scale of
nonlinear features. Finally, we need $L_z$ to be many $H$ to capture
significant $e^{z/2H}$ growth and separate the linear and nonlinear
amplitudes of $u_z$.

Fixing $L_z = 10H$ to give $\sim e^4$ amplitude growth between the
damping zones, this implies we require separation of scales $1 \ll
k_{0z}H \ll \frac{H}{L_{NL}} \ll \frac{2\pi N_z}{10}$.

Since $N_z$ is fixed by computational cost (we tried both $N_z = 512,
N_z = 1024$), we choose $k_{0z} \sim \frac{2\pi}{H}$. This is less
physically representative but gives more wavenumber space for the
nonlinear cascade.
